Concurrent Programming in Harmony

### 2021-11-15 book/paper.tex
Latest

@@

\index{ring buffer}%
A \emph{bounded buffer} (aka \emph{ring buffer}) is
a queue with the usual \texttt{put/get} interface,
but implemented using a circular buffer
\index{circular buffer}%
of a certain length and two pointers:
the \emph{tail} points where new items are enqueued and the \emph{head}
points where items are dequeued.
If the buffer is full, the enqueuer must wait; if the buffer is empty, the
dequeuer must wait.
This problem is known as the ``Producer/Consumer Problem'' and was
proposed by Dijkstra~\cite{EWD329}.
Multiple producers and multiple consumers may all share
the same bounded buffer.

@@

>>>>

\begin{figure}
\begin{code}
\harmonysource{boundedbuffer}
\end{code}
\caption{\harmonylink{code/boundedbuffer.hny} Bounded buffer specification.}
\label{fig:boundedbuffer}
\end{figure}

@@

\index{ring buffer}%
A \emph{bounded buffer} (aka \emph{ring buffer}) is
a queue with the usual \texttt{put/get} interface,
but implemented using a circular buffer
\index{circular buffer}%
of a certain length and two pointers:
the \emph{tail} points where new items are enqueued and the \emph{head}
points where items are dequeued.
If the buffer is full, the enqueuer must wait; if the buffer is empty, the
dequeuer must wait.
\autoref{fig:boundedbuffer} specifies a bounded buffer.
Coming up with a good implementation is known as the
``Producer/Consumer Problem'' and was
proposed by Dijkstra~\cite{EWD329}.
Multiple producers and multiple consumers may all share
the same bounded buffer.

@@

\index{flow control}%
Unlike an ordinary queue, where queues can grow arbitrarily, bounded buffers
provide \emph{flow control}: if the consumer runs faster than the producer
(or producers), it will automatically block until there are new requests.
Similarly, if the producers add requests at a rate that is higher than the
consumers can deal with, the producers are blocked.  While a buffer of size 1
already provides those properties, a larger buffer is able to deal with short
spikes without blocking anybody.
